Title

### 2020-06-28 paper2/paper2.tex
Created paper2 and added first version of neural network chapter

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{url}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Network Data Collection for Artifical Intelligence
}

\author{\IEEEauthorblockN{Sebastian Lipp}
\IEEEauthorblockA{\textit{IT Security} \\
\textit{FH Technikum Wien}\\
Vienna, Austria \\
cs19m032@technikum-wien.at}
\and
\IEEEauthorblockN{Damir Marijanovic}
\IEEEauthorblockA{\textit{IT Security} \\
\textit{FH Technikum Wien}\\
Vienna, Austria \\
cs19m031@technikum-wien.at}
\and
\IEEEauthorblockN{Boris Stampf}
\IEEEauthorblockA{\textit{IT Security} \\
\textit{FH Technikum Wien}\\
Vienna, Austria \\
cs19m006@technikum-wien.at}
}

@@

\begin{abstract}
Collecting real network data for further processing is an important topic for the training of neural networks which filter suspicious from ordinary traffic. This paper focuses mainly on research in the areas of network simulation, network recording and data extraction which should serve as input for the following papers.
\end{abstract}

\begin{IEEEkeywords}
artifical intelligence, neural networks, computer networks, network security, data collection
\end{IEEEkeywords}

\section{Introduction}
Neural networks which filter suspicious network traffic need a training data set of high quality. This data set could be produced from real network traffic but this can be difficult to implement. Another option considers network simulators which can be easily configured depending on the selected scenarios.

This paper is split in three chapters. The first chapter compares available network simulators and their software interfaces and explores possibilities for their automated setup based on the scenario and parameters. The second chapter focuses on tools to record the network traffic. In the third chapter methods of feature extraction and label definition are covered. This information serves as foundation for the further implementation of a system for automated data collection for artifical intelligence.

At this point all our considerations are based on theoretical research  without practical implementation which will come to life in further work. Therefore this paper sets a direction and foundation for the next steps and shows possible solutions on a high level.

\section{Simulation}

Before network simulators can be compared the requirements have to be defined. Because the produced network traffic should be close to real traffic it would be advisable to choose a network simulator which works with the simulation of real network equipment and can also integrate real hardware or virtual machines into the simulation which is necessary to redirect the network traffic to the recorder and data extractor. This is important to consider since other simulators like NS3 provide discrete-event network simulation which is a needless complexity overhead for this kind of application. It would be desirable if they are affordable and provide an easy-to-use interface to create and control specific network topologies through a programming language like Python to automate their creation. Moreover they should be resource-efficient and easy to deploy. The provisioning of the simulation hosts has to be simple and also scalable to be prepared for simulating denial-of-service attacks.

The next sections cover possible network simulators, system environments and the provisioning of the infrastructure.

\subsection{Comparing common network simulators}
There are many solutions available on the market to simulate networks. Popular network simulators which possibly fulfill the requirements above are for example VIRL, GNS3 or EVE-NG. This section focuses on the comparison of these simulators and the selection of the best solution for our use case.  \cite{b1}

\begin{figure}[htbp]
\centerline{\includegraphics[scale=0.55]{virl.png}}
\caption{Sample network topology in VIRL \cite{b2}}
\label{virl}
\end{figure}

VIRL (Cisco Virtual Internet Routing Lab) is a network modeling and simulation environment provided by Cisco which offers the ability to use simulation models of real Cisco platforms, integrate virtual machines and connect real networks. Furthermore it provides a RESTful API to configure networks without a graphical interface, sufficient documentation and scalability as well as easy deployment through prepared images. It is not for free and only supports VMware or bare-metal installs. Simulation models of other vendors are not supported. Tests in virtual machines unveiled a good responsiveness but a long topology loading time with high CPU load. Fig.~\ref{virl} shows a sample network topology created via the graphical user interface of VIRL.  \cite{b1} \cite{b3}

@@

\begin{figure}[htbp]
\centerline{\includegraphics[scale=0.13]{gns3.png}}
\caption{Sample network topology in GNS3 \cite{b4}}
\label{gns3}
\end{figure}

GNS3 (Graphical Network Simulator 3) is a very popular free open source software to emulate and test networks. It offers the integration of virtual machines and real networks and supports models of various vendors but does not offer pre-installed device images due to license issues. It also provides a RESTful API to set up network topologies, a big community and scalability. The software can be easily installed through packages on all operating systems and supports all hypervisors as well Docker containers. Performance tests showed a long topology loading time with high CPU load but a good response through the user interface. Fig.~\ref{gns3} shows a sample network topology created with GNS3.  \cite{b1} \cite{b5}

\begin{figure}[htbp]
\centerline{\includegraphics[scale=0.19]{eve-ng.png}}
\caption{Sample network topology in EVE-NG \cite{b6}}
\label{eve-ng}
\end{figure}

EVE-NG (Emulated Virtual Environment - Next Gen) is a network emulator which comes in a free and a professional edition and offers the interaction with real networks. It supports multiple vendors but does not provide device images. Advanced features like Docker, configuration management or Wireshark integration only come with the purchase version. Moreover a RESTful API and images for the installation in virtual machines are available as well as a reasonable documentation. The software is scalable and performs well in virtual machines with a short topology loading time and provides an intuitive interface as shown in Fig.~\ref{eve-ng}.  \cite{b1} \cite{b7}

All solutions are very similar in their functionalities but distinguish in their vendor support, performance, documentation and price. VIRL only supports Cisco devices but comes with device images like GNS3 and EVE-NG support various vendors but do not ship images. But many proprietary and open source images are available on the internet for free. VIRL and EVE-NG are propriatary like GNS3 is free and open source. EVE-NG performs better in virtual machines than VIRL and GNS3.

Regarding these factors GNS3 fullfills our requirements the most because it is free, supports all vendors and comes with a big community and documentation. Our second option could be EVE-NG in the professional edition which also supports all vendors.

\subsection{Selecting the deployment environment}

Next steps include the integration of the selected network simulator in a scalable environment which can be set up easily.

\begin{figure}[htbp]
\centerline{\includegraphics[scale=0.5]{docker_vm.png}}
\caption{Virtual machines vs. Containers \cite{b8}}
\label{docker-vm}
\end{figure}

Hypervisors like VirtualBox or VMware provide the possibility to set up full virtual computers on a physical host (see Fig.~\ref{docker-vm}a) while Docker containers are encapsuled sandboxes on an existing operating system (see Fig.~\ref{docker-vm}b). It is clear that containers are more resource efficent but in virtual machines the systems are better isolated. In our case we could use both solutions because the network simulators are also capable to run in Docker containers. If we focus on resource efficency and scalability also Kubernetes comes in our minds which is an orchestrator for deploying Docker containers on multiple nodes but it is a more complex configuration overhead and needs certain amount of hardware to unfold its potential also for simple scenarios. Because we want an affordable solution which can be used in most of the cases also on our computers with some exceptions, we could use a tool which can create virtual machines as well as Docker containers to be ready for the future. Regarding the corner cases servers or a Kubernetes cluster could be rented temporarily on cloud platforms like AWS to support high-load scenarios e.g. big denial-of-service attacks. \cite{b8}

A popular solution to bootstrap development environments in virtual machines is called Vagrant which provides lightweight base images and an interface to automate the configuration of the virtual machines. Moreover it can be connected to provisioning tools to further configure host systems. Vagrant supports VirtualBox or VMware but also Docker and can be extended to support cloud services like AWS or even to install a whole Kubernetes cluster. It gives us the power to start with simple configurations and scale them up to more sophisticated ones using containerization and complex orchestrators. \cite{b9} \cite{b10}

This means we would be ready to create system environments on our host or in the cloud and install base systems on them. The next section focuses on the automated setup of the network simulator, recorder and data extractor within virtual machines or Docker containers as well as the configuration of network topologies based on the selected scenarios and parameters to produce, record and extract network traffic for further processing in neural networks to improve their strike rate.

\subsection{Provisioning of the infrastructure}

Vagrant supports multiple provisioners like Ansible, Terraform or simple bash scripts to configure the host systems and can also be set up through them. For the beginning we could use Ansible through Vagrant to configure the host system. Vagrant can set up multiple virtual machines or Docker containers on one host. If we would need to set up machines on multiple hosts we could come back to tools like Terraform for example. \cite{b9} \cite{b10} \cite{b11}

If all the possible network scenarios are defined these network topologies have to be deployed through our deployment system. This means we could define the basic host setup and the network configurations for all the scenarios in Ansible templates which could be later adjusted via command line parameters. The topology can then be deployed by executing a shell command. This includes the deployment of the network recorder and the data extractor. In other words one or multiple scenarios can be started and connected to one recorder and data extractor to combine simultaneous attacks and everything will be deployed through a shell command with static configuration in templates and dynamic configuration with command line options. It will be also possible to integrate real hardware into the simulation. \cite{b9}

\begin{figure}[htbp]
\centerline{\includegraphics[scale=0.28]{design_flow.png}}
\caption{Data collection workflow}
\label{design-flow}
\end{figure}

Fig.~\ref{design-flow} shows a possible simulation workflow. The input parameters are the type of scenario(s) and the respective settings which deploys hosts including the base systems and the simulators applying the network topologies. Moreover the network recorder and data extractor could be deployed to gather the network traffic and extract the features to create data sets which could be collected in a central database.

A real test case would then consist of multiple scenarios of attacks and usual network traffic recorded, extracted and collected in a central place. The next chapter focuses on recording the generated data.

\section{Recording}

Before the network traffic can be analysed, it has to be recorded first. Therefore, this chapter will give an overview of the available third-party software which can be used to record all the incoming network traffic. The functional requirements for a tool are chiefly the pace of handling the network traffic. All of these tools are client-based tools. This means that either the network target for the traffic has to be a client or the e.g. switch has to clone a port with all the incoming and outgoing traffic and direct all the data to a client running one or more of these tools. In some cases, analysing of network traffic can take much longer than the data is sent or received by the network which means, that all the data need to be buffered or stored. In this case a packet-to-disk recorder is needed. This device can capture large amount of network traffic with a data rate of up to 40 Gbit/s and write it into files.

\subsection{pcap Library}

Every tools needs APIs\footnote{Application Programming Interface} to capture all the traffic sent through the network interface. Therefore, libraries were made for both Unix-like and Windows based operating systems. While \textit{libpcap} is used for Unix-like operating systems, a Windows system needs the library called \textit{WinPcap}. In 2013 the last release (4.1.3) of \textit{WinPcap} was published. Since then \textit{Npcap} acts as a substitute\cite{winpcaporg}. With a pcap library, software has the ability to tap all the network traffic from the selected interface. Based on the possibility to include the library in \textbf{C} and \textbf{C++}, \textit{pcap} can be used very easily in software. With the release of \textit{libpcap} 1.0.0, the zero-copy mechanism was implemented which allows the software to read the network traffic directly without copying the data from kernel space to user space.

\subsection{Network Recording Tools}

In this section we will itemise five well known and useful network recording tools. Some of the tools are console based tools, others have a GUI, or both.

\subsection*{tcpdump}

Especially in the world of Unix based operating systems, \textit{tcpdump} is one of the best known tools for monitoring and sniffing network traffic. \textit{tcpdump} is an exclusively command line based tool. This property makes \textit{tcpdump} a tool for advanced command line users only. But once you get familiar with the commands and options of \textit{tcpdump}, it will get a  a very efficient and powerful tool for many applications. \textit{tcpdump} uses the \textit{libpcap} network library of Unix based operating systems for capturing network traffic.

\begin{figure}[htbp]
\centerline{\includegraphics[scale=0.4]{tcpdump.png}}
\caption{tcpdump Network Capture}
\label{tcpdump}
\end{figure}

With the wide set of options, \textit{tcpdump} is a pretty agile tool for dumping and analysing network traffic. A few of the options are listed below \cite{tcpdumporg}.

\begin{itemize}
\item \textbf{-i eth0} the traffic from the interface \textit{eth0} is captured
\item \textbf{-w example.dump} all the output is written into a declared file
\item \textbf{-e} the ethernet header will recorded too
\item \textbf{-r example.dump} parsing the file \textit{example.dump} into tcpdump
\end{itemize}

If \textit{tcpdump} is compiled with cryptography, it is able to decrypt IPsec ESP packets in real time \cite{goyal2017}.
For the Windows enthusiasts a Windows derivative of \textit{tcpdump} is available. It is called \textit{WinDump}. \textit{WinDump} uses the \textit{WinPcap} library, which provides the same API for Windows operating systems as \textit{libpcap} for Unix-like systems. This means, network traffic captured, e.g. on a Ubuntu system can be analysed on a Windows 10 Client afterwards.

\subsection*{Wireshark}

\textit{Wireshark} is for good reason the best known cross-platform networking tool available. In 2006 \textit{Wireshark} was started as a fork of the software called \textit{Ethereal} by the Wireshark-Community due to trademark issues \cite{asrodiapatel2012}.  The performance of this tool is powerful and the set of functions is neatly arranged. By using the GTK toolkit, \textit{Wireshark} has implemented a GUI \footnote{Graphical User Interface}\cite{suribatra2012}. The GUI is separated typically into three segments: the traffic overview, and the details and raw data of the selected packet.

\begin{figure}[htbp]
\centerline{\includegraphics[scale=0.48]{wireshark.png}}
\caption{Wireshark Network Capture}
\label{wireshark}
\end{figure}

A good asset of \textit{Wireshark} is its available set of filters for capturing and displaying network data. With the help of plugins, \textit{Wireshark} can even handle with new or dedicated protocols. Another big asset of \textit{Wireshark} is its capability of capturing raw USB traffic from an USB port. By highlighting the traffic with different colours as shown in Fig.~\ref{wireshark} it's easier to scan for required traffic.

\subsection*{netsniff-ng}

\textit{netsniff-ng} is a another Unix based tool for network analysis. \textit{netsniff-ng} was the first tool which had the ability to analyse the whole traffic with a zero-copy procedure, which means that \textit{netsniff-ng} uses the same data as the kernel does. Based on this functional principle, \textit{netsniff-ng} is one of the best performing tools for network sniffing. \textit{netsniff-ng} is like \textit{tcpdump} exclusively an console based software. The set of options contains 38 capabilities on the latest version 0.6.4 of \textit{netsniff-ng}.

\subsection*{Capsa}

The fourth tool is called \textit{Capsa} from \textit{Colasoft} founded 2001 in United States. \textit{Capsa} is suitable for real-time network data capturing both LAN and Wireless Interfaces. A big advantage of \textit{Capsa} is the ability to evaluate the network traffic into graphical output like diagrams. This feature can help identifying the source of undesirable network traffic quickly by a human. The free version of \textit{Capsa} is heavily restricted regarding available functions.

\begin{figure}[htbp]
\centerline{\includegraphics[scale=0.30]{capsa.png}}
\caption{Graphical Evaluation of Network Traffic with Capsa \cite{firewallcx}}
\label{capsa}
\end{figure}

\subsection*{EtherApe}

Yet another tool for network monitoring is called \textit{EtherApe}. \textit{EtherApe} is an Unix based software with a meaningful graphic interface and graphical data output of network traffic. With the help of graphical illustration anomalies like a DDoS attack can be detected easily as shown in Fig.~\ref{etherape}. Every host is shown as a node in circle, every connection is shown as link between two nodes and every protocol has an own colour for the link. With higher network traffic between two hosts the link between nodes is shown larger.

\begin{figure}[htbp]
\centerline{\includegraphics[scale=0.37]{etherape.png}}
\caption{Graphic Representation of Network Traffic \cite{etherapewiki}}
\label{etherape}
\end{figure}

\section{Data Extraction}

Datasets for intrusion detection are often based on network packet data captured by sniffers, which can be stored in pcap files (as done in the UNSW-NB15 \cite{Nb2015} and CSE-CIC-IDS2018 \cite{Ids2018} datasets). Before a machine learning algorithm can be employed, features have to be extracted from the captured data. A wide range of different features is described in literature. For the sake of clarity, these features are divided into groups which are described in the following subsections.

There also exists a large number of techniques for dimensionality reduction. The 40 features of the NSL-KDD data set could be reduced to 30 or 16 features, with no or low degradation in detection power. The benefits of a reduced feature set are faster processing, less memory consumption, and in some cases even higher detection precision, because some classifiers have problems with irrelevant or correlated features. This could be achieved by combining the following methods \cite{vasquez2015}:
\begin{itemize}
	\item Weight by maximum relevance
	\item Minimum redundancy maximum relevance
	\item Significance analysis for microarrays
	\item Least absolute selection and shrinkage operator
	\item Stability selection
	\item Stepwise regression
\end{itemize}

Other methods which can be used for that purpose are based on association rule mining (ARM) \cite{moustafa2015}, sequential forward selection (SFS) \cite{dehghani2010} \cite{xu2015}, or the T-test \cite{sopuru2019},

When selecting the features, their limitations should be considered as well. Some features can easily be spoofed by attackers, while others are unable to detect encrypted malware traffic or cause performance problems. The development of new evasion techniques (e.g., disguising malicious traffic in widely used protocols such as HTTP) make the detection even harder \cite{celik2015}.

Another aspect which has been examined is the stability of features in terms of time (year 2000/2001 vs. 2008) and location (Internet core vs. Internet edge). In this regard, the packet size can be considered as stable feature \cite{este2009}. The suitability of features also depends on the type of network which is analyzed. For example, some industrial control system networks are relatively static regarding IP address allocation, the number of open connections and the ports, protocols and settings they use. Therefore, features based on these attributes may provide good results in that particular environment \cite{mantere2013}.

\subsection{Flow statistical features}

A common approach is to group packets into flows, and then calculate statistical features based on these flows. Generally, a flow is a group of packets with the same source and destination IP addresses and ports, and the same protocol. However, a flow can also be terminated by a TCP FIN packet or a timeout \cite{Ids2018}. Statistical features of network flows include the following:
\begin{itemize}
	\item Flow duration \cite{Nb2015} \cite{Ids2018}
	\item Time between two flows or packets \cite{Ids2018}
	\item Time between SYN and SYN/ACK, and between SYN/ACK and ACK packets during TCP connection setup \cite{Nb2015}
	\item Active time before a flow becomes idle and vice versa \cite{Ids2018}
	\item Jitter \cite{Nb2015}
	\item Number of packets in total, with at least 1 byte of payload, or with certain protocol flags \cite{Ids2018} \cite{waizumi2007} \cite{celik2015}
	\item Number of fragmented packets \cite{waizumi2007}
	\item Number of retransmitted or dropped packets \cite{Nb2015}
	\item Size of packets, payload, headers or segments \cite{Ids2018} \cite{celik2015}
	\item Number of bytes/packets transferred per second, in bulk, or in the initial TCP window \cite{Ids2018} \cite{celik2015}
	\item Number of RTT samples \cite{celik2015}
	\item Time to live value \cite{Nb2015}
	\item TCP window advertisement value \cite{Nb2015}
	\item TCP base sequence number \cite{Nb2015}
	\item Ratio between download and upload \cite{Ids2018}
	\item Ratio between maximum and minimum packet size \cite{celik2015}
	\item Number of flows with the current flow's IP address or port number \cite{waizumi2007} \cite{Nb2015}
\end{itemize}
Most of these features can be derived by statistical calculations such as total, minimum, maximum, mean, median, standard deviation, variance, skewness and kurtosis \cite{Ids2018} \cite{celik2015, xu2015}. Also, the features can be extracted from bidirectional flows, or separately for forward and backward directions \cite{Ids2018}.

The open-source Java software CICFlowMeter can be used to extract features of network flows (which end either by TCP FIN packet, or a configurable timeout) from pcap files. It allows to select from a list of available features, but can also be extended to support new features \cite{cicflowmeter}, and has therefore been used in various applications \cite{Ids2018} \cite{lashkari2017} \cite{sopuru2019}.

\begin{figure}[htpb]
\centerline{\includegraphics[scale=0.35]{hast-ids.png}}
\caption{Feature learning process of HAST-IDS \cite{wang2018}}
\label{hast-ids}
\end{figure}

The hierarchical spatial-temporal feature-based intrusion detection system (HAST-IDS), which is shown in fig.~\ref{hast-ids}, also extracts flow-based features. However, it first transforms the bytes of network flows (HAST-I) or packets (HAST-II) into images using one-hot encoding (OHE), and then uses deep convolutional neural networks (CNNs) and long short-term memory (LSTM) networks to learn features based on these images \cite{wang2018}.

\subsection{Timeslot features}

>>>>

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Title}

\author{\IEEEauthorblockN{Boris Stampf}
\IEEEauthorblockA{\textit{IT Security} \\
\textit{FH Technikum Wien}\\
Vienna, Austria\\
cs19m006@technikum-wien.at}
\and
\IEEEauthorblockN{Maximilian Wech}
\IEEEauthorblockA{\textit{IT Security} \\
\textit{FH Technikum Wien}\\
Vienna, Austria \\
cs19m020@technikum-wien.at}
}

@@

\begin{abstract}
@TBA
\end{abstract}

\begin{IEEEkeywords}
artificial intelligence, neural networks, computer networks, network security
\end{IEEEkeywords}

\section{Introduction}
@Boris

\section{Feature Extraction}
@Boris

\section{Neural Network}
The created dataset will be used in the further course of this paper to build a neural network. With this it should be possible to detect anomalies in network traffic (e. g. DoS attacks) but also harmless data packets. This is a continuation of  \cite{max1}, in which an attempt was made to create a neural network for intrusion detection on the basis of the publicly available CSE-CIC-IDS2018 dataset  \cite{max2}. The knowledge gained at that time is used, the source codes are reused accordingly and slightly adapted to the new dataset. In principle, the performance of neural networks over different datasets is to be checked. The following scenarios will be considered:
\smallskip

Scenario 1) What performance can be achieved with a neural network when it is trained and tested with the dataset created in this work? 
\smallskip

Scenario 2) How well does the neural network created in Scenario 1 work when tested with the CSE-CIC-IDS2018 dataset  \cite{max2}? 
\smallskip

Scenario 3) How well does the neural network trained in \cite{max1} work when tested with the dataset created in this work?
\smallskip

It is suspected that there are differences in the intrusion detection datasets because, for example, cyber attacks are not always executed or recorded in the same way or the network infrastructure is structured differently. The aim is to determine how sensitive neural networks react to this and are thus dependent on the data with which they have been trained. This chapter is designed in such a way that the various steps for building up the neural network are first described and then the results obtained are examined in detail.

\subsection{Setup}
For the practical implementation, a notebook with 16GB RAM and a 2.7 GHz dual-core processor was used. This hardware was sufficient for the fulfillment of the task. The training of the neural network took about 4 hours. PyCharm Community Edition 2020.1 was used as a Python development environment on Windows 10. Packages were basically the same as in \cite{max1}. The Seaborn package 0.10.1 was also used to create heatmaps.

\subsection{Preprocessing}
This process is based on the one from\cite{max1}. This involves preparing the created dataset for training the neural network accordingly. A difference compared to \cite{max1} is that this time no headers were in the middle of the created CSV file. For this reason, the "skip\_rows" parameter did not have to be used in the "read\_csv" pandas function. In addition, identical entries in the dataset (duplicates) are also removed and a corresponding data type is specified for each feature. The conversion of the labels into a numeric format also takes place. At this point it should be noted that due to time constraints, not all types of attack from \cite{max2} could be reproduced in this work. This can be seen from the fact that some (numerical) labels are not represented in the created dataset. In the context of future work, the missing types of attack can be added. It is important that the labels in the created dataset must have the exact same numerical representation as in \cite{max1}. This means, for example, that benign network traffic must be labeled with "0" again. Furthermore, no types of attack were implemented in the context of this work that are not available in the CSE-CIC-IDS2018 dataset \cite{max2}. This is because the neural network from \cite{max1} could not recognize them and thus no performance comparison would be possible. For this reason, attacks from the CSE-CIC-IDS2018 dataset \cite{max2} were selectively reproduced. It should also be noted that in \cite{max1} some attacks of the publicly available dataset were not included.

\smallskip
The following Table \ref{table:overviewdatasets} provides an overview of both datasets:
\begin{table}[htbp]
\caption{Overview of the two datasets } 
\centering
\begin{tabular}{ | c | c | c | }
\hline
Label & \multicolumn{1}{|p{2cm}|}{\centering Number of data \\ points in the dataset  \\ created in this work}  & \multicolumn{1}{|p{2cm}|}{\centering Number of data \\ points in the \\  CSE-CIC-IDS2018 dataset \cite{max2}}\\
\hline
benign (0) & 63.264 & 10.180.908 \\
\hline
ddos attacks-loic-http (1) & Not available & 575.364 \\
\hline
ddos attack-hoic (2) & 193.729	& 198.861 \\
\hline
dos attacks-hulk (3) & Not available & 145.199 \\
\hline
bot (4) & Not available & 144.535 \\
\hline
ssh-bruteforce (5) & 29.182 & 94.048 \\
\hline
dos-goldeneye (6) & 7.368.370	 & 41.406 \\
\hline
dos attacks-slowloris (7) & Not available & 9.908 \\
\hline
ddos attack-loic-udp (8) & Not available & 1.730 \\
\hline
\end{tabular}
\label{table:overviewdatasets}
\end{table}

@@

While the CSE-CIC-IDS2018 dataset \cite{max2} has a total of 10,182,638 data points (6.41GB distributed over multiple CSV files), the dataset created in this work consists of 7,654,545 data points (4.1GB, only one CSV file). When looking at Table 1, it is noticeable that both datasets are unbalanced and there is one label for which comparatively many data points exist (dataset created in this work: "dos-goldeneye", CSE-CIC-IDS2018 dataset: "benign").
The preprocessing process of the final dataset takes about 10 minutes. This is comparatively faster than in \cite{max1}, which could be due to the smaller file size and the fact that the current dataset consists of only one CSV file (there is no need to merge several files). Finally, the dataset was divided into a training and a test set at a ratio of 80:20. While the former is used to train the neural network, the latter is used to determine the performance of the neural network (with data not used for training).

\subsection{Training the neural network}
After the preprocessing of the data has been completed, the next task is to train and test the neural network. The architecture recommended in \cite{max1} (three hidden layers with 69 neurons each) and the hyperparameters (Learn rate: 0. 0001, Optimizer: adam, Weight initialization: normal, Activation function: relu, Number of neurons per layer: 69, Epochs: 850, Batch size: 256) are reused. In addition, "MinMaxScale" was used to normalize the data and a weight value of 500 was set for benign" traffic. At this point it is worth mentioning that attempts have been made to adjust individual parameters such as the learn rate or the weight value of benign traffic in order to achieve a better result. However, this has not had any positive effects. Future work could deal with this in more detail and try to find optimal hyperparameters by an extensive gridsearch. Of course, this would require appropriate hardware resources.

\smallskip
Training of the neural network with the training set took 4 hours and was stopped after 129 epochs. As in \cite{max1}, "EarlyStopping" with a value of 50 epochs was used. This means that after 50 epochs without improvement, the learning process is terminated to avoid overfitting.

\subsection{Results}
In the further course of this chapter, the aim is to present and evaluate the results achieved. To determine how good the classifications of machine learning algorithms are, an appropriate measure is required. The following metrics are relevant \cite{max3}:
\smallskip

Accuracy: How likely is the model to make a correct prediction?
\smallskip

Precision for class A: How likely is the classification to be correct if the model predicts class A?
\smallskip

Recall for class A: How much data of a class A can the model correctly predict?
\smallskip

Both the dataset created in this work and the CSE-CIC-IDS2018 dataset \cite{max2} are unbalanced (different amounts of data per class). For this reason, the use of Accuracy does not seem to be appropriate. If a very large amount of data were available for a class A and not for classes B and C, then the accuracy would be very dependent on class A. Instead, it is advisable to graphically display the classification results of the neural network in a confusion matrix and to calculate precision and recall respectively. The following text analyses the results of the three scenarios in the introduction:
Figure  \ref{fig:cm1} shows the result of the first scenario in which a neural network was trained and tested with the dataset created in this work. The classification results are largely good. However, it is noticeable, that only 62.5\% of all data of the "ssh-bruteforce" attack type can be correctly classified (comparatively low recall). The precision value for "benign" traffic is also comparatively lower, which means that it is only 86\% correct if the neural network predicts this class. All other precision and recall values are well over 95\%.

\begin{figure}[htbp]  
\centerline{\includegraphics[scale=0.65]{NeuesModellNeueDaten.png}}
\caption{Confusion Matrix for Szenario 1}
\label{fig:cm1}
\end{figure}

Next, the neural network trained in Scenario 1 was reused and tested with the CSE-CIC-IDS2018 dataset \cite{max2}. There are many more classes (attack types) in this dataset than in the one used to train the neural network. Thus, the model would not be able to recognize these classes. For this reason, it is necessary to reduce the CSE-CIC-IDS2018 dataset \cite{max2} to the known four classes ("benign", "bot", "ssh-bruteforce" and "dos-goldeneye"). The reduced dataset is then used to test the neural network. The Confusion Matrix in Figure  \ref{fig:cm2} shows that the classification result is anything but good. On the positive side, the precision for benign traffic is very high, so if the model predicts this class, the prediction is 99.4\% correct. However, only 24.2\% of all data in this class can be correctly predicted. All other precision and recall values are or move around 0\%. It is worth mentioning that the "ssh-bruteforce" attack cannot be recognized correctly once. Scenario 1 already indicated a worse classification result for this class. Both "bot"- and "dos-goldeneye" attacks could be detected at least a few times, but to a very small extent. Overall, this suggests that there may be differences in the two datasets that significantly affect the classification outcome of the neural network. Thus, there is a suspicion that the machine learning model is overfitting.

\begin{figure}[htbp]  
\centerline{\includegraphics[scale=0.65]{NeuesModellAlteDaten.png}}
\caption{Confusion Matrix for Szenario 2}
\label{fig:cm2}
\end{figure}

In the third scenario, the neural network from \cite{max1}, which was trained with CSE-CIC-IDS2018 Dataset \cite{max2}, is now tested with the dataset of this work. This continues the low classification success (see Figure \ref{fig:cm3}). In the class "benign", 99.8\% of all data are correctly predicted (recall), but the precision is noticeable low (0. 008\%). This means that the neural network recognizes many data points as "benign", but these are mostly other classes. Conversely, this is the case with "bot", where only few data of this class are correctly recognized, but whenever the neural network recognizes this class, it is a correct prediction. All other precision and recall values are or move around 0\%. It is particularly noticeable that "bot"-, "ssh-bruteforce"- and "dos-goldeneye" attacks are mostly classified as benign traffic. Furthermore, it can be seen that "ssh-bruteforce"- and "dos-goldeneye" attacks cannot be correctly detected once; bot attacks can be correctly predicted nine times in total. This makes it clear how weak the classifications of the neural network trained in \cite{max1} are with the dataset created in this work. Also worth mentioning is that for some data points the classes "ddos attacks-loic-http", "dos attacks-hulk" and "dos attacks-slowloris" are predicted. These are not available in the dataset created in this work, but in the CSE-CIC-IDS2018 dataset \cite{max2}, with which the neural network was trained. For this reason, this is not surprising.

\begin{figure}[htbp]  
\centerline{\includegraphics[scale=0.65]{AltesModellNeueDaten.png}}
\caption{Confusion Matrix for Szenario 3}
\label{fig:cm3}
\end{figure}

Scenarios 2 and 3 show that if in this context a neural network is tested with a different dataset than the one it was trained with, then the classification performance is poor. There can be many reasons for this. The suspicion, however, is that there are differences in the datasets and that this affects the predictions of the neural networks. One approach is to perform a correlation analysis. This makes it possible to determine the statistical relationship between two features (variables) of a dataset. The correlation coefficient is to be calculated, which indicates the strength of the correlation in the interval [-1.1]. A positive correlation means the larger the variable A, the larger the variable B. The opposite is the negative correlation, in which the larger the variable A is, the smaller the variable B. If the correlation coefficient is 0, this is called a neutral correlation and there is no relationship between the variables \cite{max4}. The idea is to determine the correlation between the features (variables) for each dataset individually. From this, graphical representations (heatmaps \cite{max5}) can be created in order to be able to compare whether correlation differences can be detected in the datasets. Since the number of features is high in both datasets, it makes sense to use only a subset of the features for the analysis. Thus, the correlation between the first 36 features is determined for both datasets. Figure \ref{fig:hm1} and Figure \ref{fig:hm2} each show the results for the two datasets. Clear differences are recognizable. It is worth mentioning that Figure \ref{fig:hm1} shows numerous negative correlations, which are not at all or hardly recognizable in Figure \ref{fig:hm2}. It is also noticeable that in Figure \ref{fig:hm2}, white lines occur, meaning that a feature in the dataset created in this work always has a value of zero. This shows that there are differences in the datasets.

\begin{figure}[htbp]  
\centerline{\includegraphics[scale=0.8]{Heatmap-AlteDaten-Filtered-0-35.png}}
\caption{Heatmap of the first 36 features of the CSE-CIC-IDS2018 Dataset}
\label{fig:hm1}
\end{figure}

\begin{figure}[htbp]  
\centerline{\includegraphics[scale=0.8]{Heatmap-NeueDaten-0-35.png}}
\caption{Heatmap of the first 36 features of the dataset created in this work}
\label{fig:hm2}
\end{figure}

\section{Conclusion}
In summary, the performance of neural networks in terms of intrusion detection over different datasets is rather mixed. If these machine learning algorithms are tested with the same dataset with which they were trained, the result is promising. Using another dataset for testing results in a poor classification performance. However, there is lots of potential for future research with neuronal networks in terms of intrusion detection. It seems useful to analyse in detail why the mentioned difficulties occur in the classification and which factors are relevant for the neural network. An intensive examination of the finding of optimal hyperparameters would also be desirable in the course of future work. Furthermore, it would also make sense to simplify the way classification is done. Instead, a simple classification like network traffic is “benign” or “not benign” would be conceivable. Moreover, it should be checked whether a better performance can be achieved with other machine learning algorithms.

\begin{thebibliography}{00}
\bibitem{max1} F. Ahrens, C. Cat and R. Freingruber, "A Neural-Network Approach for an Intrusion Detection System", FH Technikum, Vienna, 2019.
\bibitem{max2} I. Sharafaldin, A. H. Lashkari and A. A. Ghorbani, "Toward Generating a New Intrusion Detection Dataset and Intrusion Traffic Characterization", in 4th International Conference on Information Systems Security and Privacy (ICISSP), Portugal, 2018.
\bibitem{max3} D. Meyer, "Performance Assessment \& Tuning - Presentation slides", FH Technikum, Vienna, 2018.
\bibitem{max4} J. Brownlee, "How to Calculate Correlation Between Variables in Python", 3 June 2020. [Online]. Available: https://machinelearningmastery.com/how-to-use-correlation-to-understand-the-relationship-between-variables/. [Accessed 27 June 2020].
\bibitem{max5} S. Norena, "Finding Correlation Between Many Variables (Multidimensional Dataset) with Python", 26 April 2018. [Online]. Available: https://medium.com/@sebastiannorena/finding-correlation-between-many-variables-multidimensional-dataset-with-python-5deb3f39ffb3. [Accessed 27 June 2020].
\end{thebibliography}

\end{document}
